\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-04-23 trunk/related.tex
updates to reviews

With few exceptions, most AUV control systems have been variations of
reactive approaches.  Typically, waypoint-based pre-defined mission
designs are uploaded to the AUV; specialized code fragments called
\emph{behaviors} are designed for the specific mission and a choice of
behaviors for the mission are used on the computational stack
\cite{bellingham94}. Swapping in and out of these behaviors using
conditionals or cleverer and quantitative approaches such as
\cite{Benjamin:2004} forms the basis for adaptation and safety in the
vehicle. Where such swapping does not aid adaptation, individual
behaviors are tweaked to generate some measure of responsiveness to
the surrounding environment \cite{yanwu08} to generate interesting
observations for oceanographers, yet technically naive in terms of
scalability and systematicity. \cite{Benjamin2006} is clear about the
extant of adaptation in stating:

@@

There are costs associated with such inference based control which are 
not inconsequential. Foremost among them is the steep learning curve
that comes with the design of the domain model. With limited support
tools necessary for knowledge capture and design of the constraints
often in stylized higher-level languages (see \cite{NDDL} for an
example), it requires the modeler to think in alternative ways to do
problem solving. More importantly it is often necessary for the
modeler to be exposed to the internal representation of the planner
and how it performs search. The level of expertise is often well above
what would be expected of a typical well-rounded Computer Science
graduate student.

However this is mild criticism given the end benefits and the
scientific and engineering goals associated with adaptive and
persistent observation for marine robots whether for civil or military
applications. Given the general problem-solving nature of such
deliberative systems, there is also a substantial overall benefit for
the field of robotics. 
\fcomment{Though it woulfd  be usefull to emphasize this ...}
The cost of modeling put aduring the intial design, for example,
should be balanced against the cost of redesigning
a new mission script eate every new deployment in more ``classical''
script-based controllers.

\subsection{Evolution of AI-based Robotic Planning: A Necessary Digression}

@@

The dominant approach for building agent control systems utilize a
$3$-layered architecture (see \cite{gat98} for a well thought out
rationale), notable examples of which include \texttt{IPEM}
\cite{AmbrosIngerson88}, \texttt{ROGUE} \cite{Haigh98}, the LAAS
Architecture \cite{alami:1998p820}, the Remote Agent Experiment
\texttt{RAX} \cite{mus98} and the Autonomous Spacecraft Experiment
\texttt{ASE} \cite{chien99} (see \cite{Knight01} for a survey). Early
efforts were cognizant of if absent with, dealing with time and
resources for such experiments.\fcomment{This sentence does not make
  sence to me}

@@

The Autonomous Sciencecraft Experiment (\texttt{ASE}) \cite{chien99,
  chien03} was another autonomy demonstration in space, this time for
a more observable vehicle, the EO-1 in Earth orbit. It encapsulated a
classical 3-layered architecture like \texttt{RAX}. The fundamental
functional difference was that goals could not only be sent by a
ground segment, but also by an onboard science driver which could
encapsulate a range of Machine Learned feature detectors which could
trigger the planner. However ASE was not an integrated system; the
\texttt{CASPER} planner \cite{chien00} would generate plans which were
executed by a separate commercially available rule-based system
\cite{icl}. \texttt{CASPER} touts a \emph{continuous} paradigm for
planning; however this incremental modification simply takes the
overall mission plan and selectively plans near-term at a more
detailed level. The system also uses iterative-repair as a way to fix
plans; however this too is derived from a model of the mission and
spacecraft constraints which allow simple reconfiguration modes
towards replanning. Robust software engineering is also not an
important aspect targeted by \texttt{ASE} given the disparate models
between the planner and the rule-based executive. With such a system,
real-time updates from the environment cannot be accommodated since
the planner is a monolithic computational environment. The rate of
updates from the science drivers is roughly in the order of an earth
orbit on the EO-1 vehicle. As a result, the ASE is only an incremental
update on RAX and is not suitable for domains where complex automated
reasoning and asynchronous environmental observations need to be
factored in deliberation.

\texttt{TCA} \cite{simmons94} is another framework for robot control
and has similar motivations to \texttt{RAX} and
\texttt{ASE}. Reactivity and deliberation are also key to \texttt{TCA}
and a systematic approach to development led to its design. However it
suffers from three key weaknesses. First it has a weaker
representational framework of Task Trees which require representation
of tasks within cleanly formulated hierarchies. Constraints between
branches within a task hierarchy are allowed; timing between leaves of
different trees is however not possible. Second, its temporal
framework does not deal with flexibility; time points are fixed and
while partial orders are possible, it does deal with execution time
uncertainty. Third, while \texttt{TCA} shares the notion of using
domain expertise of different planners, it does so thru a centralized
message passing mechanism. While this allows disparate code-bases to
talk to each other thru IPC (a inter-process communication or message
passing mechanism), failure in the centralized controller is
tantamount to a system crash.

@@

Our system, \texttt{T-REX}, the focus of this chapter, is an agent
control structure formally defined as a composition of coordinated
control loops where sensing, planning and acting are resulting from 
concurrent deliberating tasks within a formal framework. \fcomment{Got
  rid of sense-plan-act reference here. \rx plan/sense while sensing/planning.}
\texttt{T-REX} was designed after carefully
evaluating lessons learned from \texttt{IDEA} and owes substantially
to the notion of partitioned problem solving from that effort. However
this partitioning is systematic and methodologically principled where
we manage the information flow within the partitioned structure to
ensure consistency in order to direct the flow of goals and
observations in a timely manner. The resulting control structure
improves scalability since many details of each controller can be
encapsulated within a single control loop.  Furthermore, partitioning
increases robustness since controller failure can be localized to
enable graceful system degradation, making this an effective
\emph{divide-and-conquer} approach to the overall control problem.

A recent variation of the key idea of such controller composition
derived from \texttt{T-REX} is \texttt{ROAR} \cite{degroote11} where
lower-level functional modules that control specific sensors are
organized within a graph hierarchy. One reason they cite such a
structure is essential is in dealing robustly to failure; an
off-nominal condition will allow rapid graph traversal to identify
alternative ways in which sensing tasks can be executed and therefore
aiding in responsiveness. However, it is unclear whether to date this
system actually deliberates and whether it is actually instantiated on
a real platform for control.

>>>>

With few exceptions, most AUV control systems have been variations of
reactive approaches.  Typically, waypoint-based pre-defined mission
designs are uploaded to the AUV; specialized code fragments called
\emph{behaviors} are designed for the specific mission and a choice of
behaviors for the mission are used on the computational stack
\cite{bellingham94}. Swapping in and out of these behaviors using
conditionals or cleverer and quantitative approaches such as
\cite{Benjamin:2004} forms the basis for adaptation and safety in the
vehicle. Where such swapping does not aid adaptation, individual
behaviors are tweaked to generate some measure of responsiveness to
the surrounding environment \cite{yanwu08} to generate interesting
observations for oceanographers, yet technically naive in terms of
scalability and systematicity. \cite{Benjamin2006} is clear about the
\kcomment{extent} of adaptation in stating:

@@

There are costs associated with such inference based control which are
not inconsequential. Foremost among them is the steep learning curve
that comes with the design of the domain model. With limited support
tools necessary for knowledge capture and design of the constraints
often in stylized higher-level languages (see \cite{NDDL} for an
example), it requires the modeler to think in alternative ways to do
problem solving. More importantly it is often necessary for the
modeler to be exposed to the internal representation of the planner
and how it performs search. The level of expertise is often well above
what would be expected of a typical well-rounded Computer Science
graduate student. Yet in our experience, the cost of model design in
such systems, is well balanced against the cost of re-engineering new
mission scripts for new deployment in more ``classical'' script-based
controllers.

However this is mild criticism given the end benefits and the
scientific and engineering goals associated with adaptive and
persistent observation for marine robots whether for civil or military
applications. Given the general problem-solving nature of such
deliberative systems, there is also a substantial overall benefit for
the field of robotics.

\subsection{Evolution of AI-based Robotic Planning: A Necessary Digression}
\label{sec:related:robotplans}

@@

The dominant approach for building agent control systems utilize a
$3$-layered architecture (see \cite{gat98} for a well thought out
rationale), notable examples of which include \texttt{IPEM}
\cite{AmbrosIngerson88}, \texttt{ROGUE} \cite{Haigh98}, the LAAS
Architecture \cite{alami:1998p820}, the Remote Agent Experiment
\texttt{RAX} \cite{mus98} and the Autonomous Spacecraft Experiment
\texttt{ASE} \cite{chien99} (see \cite{Knight01} for a survey). % Early
% efforts were cognizant of if absent with, dealing with time and
% resources for such experiments.

@@

The Autonomous Sciencecraft Experiment (\texttt{ASE}) \cite{chien99,
  chien03} was another autonomy demonstration in space, this time for
a more observable vehicle, the EO-1 in Earth orbit. It encapsulated a
classical 3-layered architecture like \texttt{RAX}. The fundamental
functional difference was that goals could not only be sent by a
ground segment, but also by an onboard science driver which could
encapsulate a range of Machine Learned feature detectors which could
trigger the planner. However ASE was not an integrated system; the
\texttt{CASPER} planner \cite{chien00} would generate plans which were
executed by a separate commercially available rule-based system
\cite{icl}. \texttt{CASPER} touts a \emph{continuous} paradigm for
planning; however this incremental modification simply takes the
overall mission plan and selectively plans near-term at a more
detailed level. The system also uses iterative-repair as a way to fix
plans; however this too is derived from a model of the mission and
spacecraft constraints which allow simple reconfiguration modes
towards replanning. Robust software engineering is also not an
important aspect targeted by \texttt{ASE} given the disparate models
between the planner and the rule-based executive. With such a system,
real-time updates from the environment cannot be accommodated since
the planner is a monolithic computational environment. The rate of
updates from the science drivers is roughly \kcomment{on} the order of
an earth orbit on the EO-1 vehicle. As a result, the ASE is only an
incremental update on RAX and is not suitable for domains where
complex automated reasoning and asynchronous environmental
observations need to be factored in deliberation.

\texttt{TCA} \cite{simmons94} is another framework for robot control
and has similar motivations to \texttt{RAX} and
\texttt{ASE}. Reactivity and deliberation are also key to \texttt{TCA}
and a systematic approach to development led to its design. However it
suffers from three key weaknesses. First it has a weaker
representational framework of Task Trees which require representation
of tasks within cleanly formulated hierarchies. Constraints between
branches within a task hierarchy are allowed; timing between leaves of
different trees is however not possible. Second, its temporal
framework does not deal with flexibility; time points are fixed and
while partial orders are possible, it does deal with execution time
uncertainty. Third, while \texttt{TCA} shares the notion of using
domain expertise of different planners, it does so thru a centralized
message passing mechanism. While this allows disparate code-bases to
talk to each other \kcomment{through} IPC (\kcomment{an} inter-process
communication or message passing mechanism), failure in the
centralized controller is tantamount to a system crash.

@@

Our system, \rxe, the focus of this chapter, is an agent control
structure formally defined as a composition of coordinated control
loops where sensing, planning and acting are resulting from concurrent
deliberating tasks within a formal framework.  \rx was designed after
carefully evaluating lessons learned from \texttt{IDEA} to which it
owes substantially the notion of partitioned problem solving. However
partitioning in \rx is systematic and methodologically principled
where we manage the information flow within the partitioned structure
to ensure consistency in order to direct the flow of goals and
observations in a timely manner. The resulting control structure
improves scalability since many details of each controller can be
encapsulated within a single control loop.  Furthermore, partitioning
increases robustness since controller failure can be localized to
enable graceful system degradation, making this an effective
\emph{divide-and-conquer} approach to the overall control problem.

A recent variation of the key idea of such controller composition
derived from \rx is \texttt{ROAR} \cite{degroote11} where lower-level
functional modules that control specific sensors are organized within
a graph hierarchy. One reason they cite such a structure is essential
is in dealing robustly to failure; an off-nominal condition will allow
rapid graph traversal to identify alternative ways in which sensing
tasks can be executed and therefore aiding in responsiveness. However,
it is unclear whether to date this system actually deliberates and
whether it is actually instantiated on a real platform for control.

### 2012-04-23 trunk/results.tex
updates to reviews

A significant driver of the applicability of our work was the growing
need for AUV missions to be driven by sensor data. This trend is
illustrated by the implementation of MBARI's \emph{Gulper}
water-samplers \cite{Bird07}, $2$ litre syringe style containers $10$
of which are encased within the mid-body section of our Dorado vehicle
as shown in Fig \ref{fig:gulper}. However, the original design and
intended use of operation continued to be dominated by a naive belief
of collecting and returning water samples as traditional ship-borne
methods often employ, namely using pre-scripted methods. With the
installation of \rx onboard the Dorado with its capability sense the
environment and to appropriately synthesize plans and replan on the
fly, more imaginative methods of sample acquisition were targeted.

@@

\begin{figure*}
\centering
\subfloat[]{\label{fig:odss-mtg}\includegraphics[width=0.3\textwidth]{figs/october-canon-meeting.png}}
\subfloat[]{\label{fig:odss-virtual}\includegraphics[width=0.33\textwidth]{figs/virtual-drifter-tool.png}}
\subfloat[]{\label{fig:odss-ipad}\includegraphics[width=0.3\textwidth]{figs/photo-ipad-2.jpg}}
\caption{\small \ref{fig:odss-mtg} shows a 2010 \can meeting with the
  \od in use. \ref{fig:odss-virtual} shows a drifter tracking tool
  used during \can planning; it used a physical model of advected
  surface currents from \texttt{CeNCOOS} \cite{cencoos} to plan AUV
  deployments. \ref{fig:odss-ipad} shows \od use on an iPad.}
\label{fig:odss2}
\end{figure*}

>>>>

A significant driver of the applicability of our work was the growing
need for AUV missions to be driven by sensor data. This trend is
illustrated by the implementation of MBARI's \emph{Gulper}
water-samplers \cite{Bird07}, $2$ litre syringe style containers $10$
of which are encased within the mid-body section of our Dorado vehicle
as shown in Fig \ref{fig:gulper}. However, the original design and
intended use of operation continued to be dominated by a naive belief
of collecting and returning water samples as traditional ship-borne
methods often employ, namely using pre-scripted methods. With the
installation of \rx onboard the Dorado with its capability sense the
environment and to appropriately synthesize plans and replan on the
fly, more imaginative methods of sample acquisition were targeted.
\kcomment{The Dorado uses two onboard computers: a main vehicle computer which
is a 244 Mhz PC/104 stack running the QNX real-time operating system,
and a 367 MHz EPIC EPX-GX500 AMD Geode stack running Linux and \rxe. A
socket-based communication protocol allows the exchange of goals and
state updates between \rx and the lower level functional layer.}

@@

\begin{figure*}
\centering
\subfloat[]{\label{fig:odss-mtg}\includegraphics[width=0.3\textwidth]{figs/october-canon-meeting.png}}
\subfloat[]{\label{fig:odss-virtual}\includegraphics[width=0.33\textwidth]{figs/virtual-drifter-tool.png}}
\subfloat[]{\label{fig:odss-ipad}\includegraphics[width=0.3\textwidth]{figs/photo-ipad-2.jpg}}
\caption{\small Fig. \ref{fig:odss-mtg} shows a 2010 \can meeting with
  the \od in use. Fig. \ref{fig:odss-virtual} shows a drifter tracking
  tool used during \can planning; it used a physical model of advected
  surface currents from \texttt{CeNCOOS} \cite{cencoos} to plan AUV
  deployments. Fig. \ref{fig:odss-ipad} shows \od use on an iPad.}
\label{fig:odss2}
\end{figure*}

### 2012-04-23 trunk/setobook.tex
added people from europa dev to acknowledgements section.
updates to reviews

Rajan and Py are funded by a block grant from the David and Lucile
Packard Foundation to MBARI. They thank their collaborators John Ryan,
Julio Harvey, Jim Bellingham and Robert Vrijenhoek at MBARI. In
addition they thank colleagues Martha del Alto, David Korsmeyer and
Steven Zornetzer at NASA Ames who made \eu available as open
source. The authors are indebted to Nicola Muscettola for his ideas
and early efforts in Constraint-based Temporal Planning at NASA Ames
which forms the basis of this work.

>>>>

Rajan and Py are funded by a block grant from the David and Lucile
Packard Foundation to MBARI. They thank their collaborators John Ryan,
Julio Harvey, Jim Bellingham and Robert Vrijenhoek at MBARI. They
thank other members of the \eu team over the years, including Conor
McGann, Ari J\'onsson, Jeremy Frank, Paul Morris, John Bresina,
Michael Iatauro, Matthew Boyce and Tristan Smith.  The authors are
grateful to Martha del Alto, David Korsmeyer and Steven Zornetzer all
at NASA Ames \kcomment{who not only made \eu available as open source
  but encouraged its use outside the agency}. Minh Bin Do (NASA Ames),
William Yew Teck Tan (NUS, Singapore) and Rishi Graham (MBARI)
provided valuable comments on early drafts; we are grateful for their
help in making this manuscript more accessible. Finally, the authors
are indebted to Nicola Muscettola \kcomment{(Google)} for his ideas
and early efforts in Constraint-based Temporal Planning while at NASA
Ames, and which forms the basis of much of this work.
